Variation of Euclidean Algorithm for finding gcd of two numbers

### 2018-01-19 latex/draft.tex
Add diagrams

The methodology for finding this greatest common divisor, was initially formulated by the famous Greek mathematician Euclid. This methodology for finding greatest common divisor of two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two natural numbers. The definition of the Euclidean Algorithm can be seen as below.
\newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y, then} \newline 
\gcd(\var{x}, \var{y}) = \gcd(\var{x} $\mod $ \var{y}, \var{y}).
\newline 
\newline An algorithm based on above definition can be found from text book [1] as follow.
%\linebreak
\\~\\
\textbf{Algorithm 1:}
\newline
\textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline
\underline {function Euclid}(\progvar{a}, \progvar{b})
\newline 
Input: Two integers \progvar{a} and \progvar{b} with \progvar{a} $\geq$ \progvar{b} $\geq$ 0 
\newline Output: gcd((\progvar{a}, \progvar{b})
\newline
\newline if \progvar{b}=0: return \progvar{a}
\newline return Euclid(\progvar{b}, \progvar{a} mod \progvar{b})
\newline 
\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator($\mod $ in short).
For two integers \var{a} and \var{b}, the working of modulus operator can be explained as follows.
For \var{a} > \var{b}:
\begin{equation} \label{eq:gcd_modulus_1}
%\> inserts space
a \> \mod \> b = r,  b > r
\end{equation}
and
\begin{equation} \label{eq:gcd_modulus_2}
a = q * b + r
\end{equation}

@@

% ---------------------------------TABBLE 1 -----------------------------------
% [t] indicates the top of the page
% {table*} used for wide table
\begin{table}[!htbp]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{Table showing equivalence class for modulo 5 on integers}
\label{table:1}
\centering
	\begin{tabular}{c c c c c c}
		... & -5 & 0 & 5 & 10 & ... \\
		... & -4 & 1 & 6 & 11 & ... \\
		... & -3 & 2 & 7 & 12 & ... \\
		... & -2 & 3 & 8 & 13 & ... \\
		... & -1 & 4 & 9 & 14 & ...
	\end{tabular}
\end{table}
% ---------------------------------TABBLE 1 -----------------------------------

And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
Symbolically it can be represented as, 7 $\equiv$ 12 (mod 5).
So when, the modulo \var{b} give same \var{r} for a set of different numbers, finding and studying the equivalence for \var{r} was enticing.

@@

If we compare the contents of table 2 and 3, we could see some common numbers and some of them differs. One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete. This is because for 
this combination of numbers we have received 6 times \var{s} > \var{r} and 2 times 
\var{s} < \var{r}. But for other combination of numbers, the scenario could be different.

@@

\section{Conclusion}
From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and it brings down the number of recursions. And also we have seen that, we have a substitute for \var{r} in the context of modulus arithmetic and that can be used within the intermediate steps of the algorithm to find gcd of two numbers. The running time of the loop evaluated to be $\log_4(b)$ number of iteration. The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands! Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash $\>$ is implemented with the help of existing \mod $\>$ operator and an additional subtraction. The running time of subtraction is linear however the running time of modular operation is quadratic in nature. So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the recursion. There are also other highly optimized algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc. for finding the gcd and this paper does not intent to provide a optimize algorithm for gcd. But the finding discussed in the paper can be clubbed together with other algorithm to make it more efficient.

>>>>

The methodology for finding this greatest common divisor, was initially formulated by the famous Greek mathematician Euclid. This methodology for finding greatest common divisor of two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two natural numbers. The definition of the Euclidean Algorithm can be seen as below.
\newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y, then} \newline 
\gcd(\var{x}, \var{y}) = \gcd(\var{x} $\mod $ \var{y}, \var{y}).
\newline 
\newline An algorithm based on the above definition can be found from text book [1] as follow.
%\linebreak
\\~\\
\textbf{Algorithm 1:}
\newline
\textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline
\underline {function Euclid}(\progvar{a}, \progvar{b})
\newline 
Input: Two integers \progvar{a} and \progvar{b} with \progvar{a} $\geq$ \progvar{b} $\geq$ 0 
\newline Output: gcd((\progvar{a}, \progvar{b})
\newline
\newline if \progvar{b}=0: return \progvar{a}
\newline return Euclid(\progvar{b}, \progvar{a} mod \progvar{b})
\newline 
\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator($\mod $ in short).
For two integers \var{a} and \var{b}, the working of modulus operator can be explained as follows.
For \var{a} > \var{b}:
\begin{equation} \label{eq:gcd_modulus_1}
%\> inserts space
a \> \mod \> b = r,  b > r
\end{equation}
and
\begin{equation} \label{eq:gcd_modulus_2}
a = q * b + r
\end{equation}

@@

% ---------------------------------TABBLE 1 -----------------------------------
% [t] indicates the top of the page
% {table*} used for wide table
\begin{table}[!htbp]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{Table showing equivalence classes of modulo 5 on integers}
\label{table:1}
\centering
	\begin{tabular}{c c c c c c}
		... & -5 & 0 & 5 & 10 & ... \\
		... & -4 & 1 & 6 & 11 & ... \\
		... & -3 & 2 & 7 & 12 & ... \\
		... & -2 & 3 & 8 & 13 & ... \\
		... & -1 & 4 & 9 & 14 & ...
	\end{tabular}
\end{table}
% ---------------------------------TABBLE 1 -----------------------------------

And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behave exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
Symbolically it can be represented as, 7 $\equiv$ 12 (mod 5).
So when, the modulo \var{b} gives same \var{r} for a set of different numbers, finding and studying the equivalence for \var{r} was enticing.

@@

If we compare the contents of table 2 and 3, we could see some common numbers and some of them differs. One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete. This is because, for 
this combination of numbers we have received 6 times \var{s} > \var{r} and 2 times 
\var{s} < \var{r}. But for other combination of numbers, the scenario could be different.

@@

\section{Conclusion}
From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and it brings down the number of iteration(or recursions). And also we have seen that, we have a substitute for \var{r} in the context of modulus arithmetic and that can be used within the intermediate steps of the algorithm to find gcd of two numbers. The running time of the loop evaluated to be $\log_4(b)$ number of iteration. The order of optimization is still under investigation and could be significant when we have integers with very very long digits may be in the order of thousands! Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash $\>$ is implemented with the help of existing \mod $\>$ operator and an additional subtraction. The running time of subtraction is linear however the running time of modular operation is quadratic in nature. So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the iteration(or recursion). There are also other highly optimized algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc. for finding the gcd and this paper does not intent to provide a optimize algorithm for gcd. But the finding discussed in the paper can be clubbed together with other algorithm to make it more efficient.
